---
title: "Quiz 1"
author: "Daniel Detore"
output: pdf_document
---
I Pledge my honor that I have abided by the Stevens Honor System.
```{r}
srs <- c(170, 129, 172, 174, 166, 169, 188, 181, 175, 164, 175, 177, 
182, 173, 179, 167, 202, 163, 160, 162)
```
# Problem 1
```{r}
boxplot(srs, horizontal=TRUE, main="x' samples boxplot")
# mean:
abline(v = mean(srs), col="red")
mean(srs)
# 86th quantile:
quantile(srs, probs=.86)
abline(v = quantile(srs, probs=.86), col="orange")
# all quartiles:
summary(srs)
abline(v = quantile(srs, probs=.25), col="yellow")
abline(v = quantile(srs, probs=.5), col="green")
abline(v = quantile(srs, probs=.75), col="blue")
```
The outliers, based on the box plot, are the min and max values 129 and 202.
```{r}
max(srs)
min(srs)
```

# Problem 2
```{r}
qqnorm(srs)
qqline(srs)
IQR(srs)
var(srs)
sd(srs)
summary(srs)
```
Based on this summary, 
the data is skewed left because the mean is smaller than the median.

# Problem 3
## i

$T_1 \sim \chi^2_n$ where $n=1$ because $T_1$ follows the form $\Sigma_{i=0}^nX_i^2$ where $X \sim \mathcal{N}(0,1)$ and all observations are independent.

This means that $P(T_1 \leq t_1) = pchisq(t_1,1)$.

## ii
According to the fundamental theorem (2):
$\Sigma_{i=1}^n \frac{(X_i-\bar{X})^2}{\sigma^2} = \frac{(n-1)S^2}{\sigma^2} \sim \chi^2_{n-1}$

This means that $P(T_2 \leq t_2) = pchisq(t_2,19)$.

## iii
The probability $P(T_2 \leq t_2)$ is larger because, as the degrees of freedom 
increase, more of the $\chi^2$ distribution's area moves to its tails. 
Since this probability is cumulative, and therefore collects most of its value 
from the distribution's left tail, it will usually be greater than the probability 
$P(T_1 \leq t_1)$.

## iv
The form of $T_3$ implies that it follows $\mathcal{N}(0,1)$. 
This is the standardization of the original distribution.

$P(|T_3| > |t_3|) = P(T_3 < -|t_3|) + 1 - P(T_3 < |t_3|)$
$= pnorm(-|t_3|,0,1) + 1 -pnorm(|t_3|,0,1)$.

$P(|T_3| > |t_3|) = P(|Z| > |t_3|)$, because $T_3 \equiv Z \sim \mathcal{N}(0,1)$,
therefore it cannot be that $P(|T_3| > |t_3|) > P(|Z| > |t_3|)$.


## v
According to a corrollary of the fundamental theorem, 
$T_4$ is of the form $\frac{N(0,1)}{\sqrt{\chi^2_n/n}} \implies T_4 \sim \mathcal{T}_{n-1}$.

$P(|T_4| > |t_4|) = P(T_4 < -|t_4|) + 1 - P(T_4 < |t_4|)$
$= pt(-|t_3|,0,1) + 1 - pt(|t_3|,0,1)$.

$P(|T_4| > |t_4|) > P(|Z| > |t_3|)$, because the $\mathcal{T}$ distribution
has much more weight in its tails, even with $df=1$, than the standard
normal distribution.


# Problem 4
In order of left-to-right then top-to-bottom:

$3X_2 +4 \sim \mathcal{N}(4,9)$

$-4X_1+3X_2+2X_3-5 \sim \mathcal{N}(-4(-1)+2(1)-5, (-4)^2(9)+3^2(1)+2^2(16)) \equiv \mathcal{N}(1, 217)$

$\equiv 3 * \chi^2_1 = \chi^2_3$

$\equiv \frac{\frac{1}{2}\chi^2_2}{\chi^2_1/1} \sim \mathcal{F}_{2,1}$

$\sim \mathcal{T}_2$

$\sim \mathcal{F}_{1,2}$  

# Problem 5
## i
$\hat{\mu} = \bar{X} = \frac{\Sigma_{i=1}^n X_i}{n}$

$\hat{\sigma^2} = S^2 = \frac{\Sigma_{i=1}^n (X_i - \bar{X})^2}{n}$

## ii

$\mathcal{T}_{n-1}$

$\chi^2_{n-1}$

# iii
It is expected that $\hat{\sigma^2} = \frac{\Sigma_{i=1}^n (X_i - \bar{X})^2}{n}$
will approximate the true $\sigma^2$ as the sample size increases.


# iii
They are not independent because you need $\bar{X} = \hat{\sigma}$ to calculate 
$S^2 = \bar{\sigma^2}$. They are correlated.

# Problem 6
It would most likely be near 0, since $x$ is random. It is incredibly unlikely for 
all $x$s to fall into a perfect linearly increasing correlation.