---
title: "q2 example"
author: "Daniel Detore"
output: pdf_document
---

```{r}
CropYield = data.frame(read.csv("./Crop yield.csv"))
CropYield$block = as.factor(CropYield$block)
CropYield$density = as.factor(CropYield$density)
CropYield$fertilizer = as.factor(CropYield$fertilizer)
```

# 1
## 1.1
$H_0:$ 'Block' and 'Density' have no association.

$H_a:$ 'Block' and 'Density' have some association.

## 1.2
```{r}
table = table(CropYield$density, CropYield$block); table
```

##1.3

```{r}
barplot(table, legend=TRUE, beside=TRUE, main='Density/Block Barplot')
```
They are of different patterns. Blocks 1 and 3 only have density 1, while blocks 2 and 4 only have density 2.

## 1.4
The value of each inner cell $e_{i,j} = \frac{o_{i,+}o_{+,j}}{o_{+,+}}$.
```{r}
o_ip <- rowSums(table)
o_pj <- colSums(table)
o_pp <- sum(table)
e <- outer(o_ip, o_pj)/o_pp
rownames(e) = c("Density 1", "Density 2")
colnames(e) = c("Block 1", "Block 2", "Block 3", "Block 4")
e
```
## 1.5
```{r}
chisq.test(table)
```
The testing statistic $$\chi^2 = \sum_{i=1}^r \sum_{j=1}^c \frac{(O_{i,j}-E_{i,j})^2}{E_{i,j}}$$ 
is observed as $$\sum_{i=1}^r \sum_{j=1}^c \frac{(o_{i,j}-e_{i,j})^2}{e_{i,j}}=96$$

$H_0$ is $\chi^2_k$ where $k=(r-1)(c-1) = (2-1)(4-1)=3$. The p-value is less than $2.2\times 10^{-16}$.

## 1.6
In this case we must reject $H_0$ because $2.2\times 10^{-16} < \alpha = 0.1$. Thus we assume $H_a$, that 'Block' and 'Density' have some association.

## 2
## 2.1
$H_0:\ \mu_1 = \mu_2 = \mu_3$

$H_a:$ mean yields due to fertilizer are not all equal.

## 2.2
```{r}
boxplot(xlab = "Fertilizer", ylab = "Yield", CropYield$yield ~ CropYield$fertilizer, beside = TRUE, main = "Fertilizer/Yield Boxplot")
```
The variance of the yields per fertilizer are varied. From fertilizer 1 through 3, the means seem to increase.

## 2.3
$$SSB = \sum_{i=1}^k n_i(\bar{X}_{i,.}-\bar{X}_{.,.})^2$$

$$SSE = \sum_{i=1}^k \sum_{j=1}^{n_i} (X_{i,j}-\bar{X}_{i,.})^2 = \sum_{n_i-1}S_i^2$$

$$MSB = \frac{SSB}{k-1}$$

$$MSE = \frac{SSE}{n-k}$$

## 2.4
```{r}
summary(aov(yield ~ fertilizer, CropYield))
```

## 2.5
Our testing statistic $F=\frac{MSB}{MSE}=\frac{\frac{SSB}{k-1}}{\frac{SSE}{n-k}}$ is observed as $f=7.86$.
The null distribution is $\mathcal{F}_{k-1,n-k}=\mathcal{F}_{3-1,96-3}=\mathcal{F}_{2,93}$ . The p-value is $7\times 10^{-4}$.

## 2.6
Since $\alpha = 0.05 > 7\times 10^{-4}$, we must reject $H_0$ and assume $H_a$, that is that the mean yields due to fertilizer are not all equal.

## 2.7
Given means per fertilizer:
```{r}
tapply(CropYield$yield, CropYield$fertilizer, mean)
```
We can check if the mean yield for fertilizer 3 is close to the mean yield across fertilizers 1 and 2, i.e. whether $-\frac{1}{2}\mu_1 - \frac{1}{2}\mu_2 + \mu_3 = 0$.

## 2.8
```{r}
contrasts(CropYield$fertilizer) = c(-1/2, -1/2, 1)
go = aov(yield ~ fertilizer, CropYield)
summary(go, split = list(fertilizer = list(`mu3 vs 1/2 mu1 + 1/2 mu2` = 1)))
```
Since the p-value $0.000258 < \alpha = 0.01$, we must reject $H_0$ and assume $H_a$, that fertilizer 3 is different than the mean yield across fertilizers 1 and 2.

## 2.9
We can do multiple comparison using Fischer's Least Significant Difference.
```{r}
library(agricolae)
comparison = LSD.test(go, "fertilizer", p.adj = "none")
comparison
```
And we can visualize the difference in standard deviation as such:
```{r}
plot(comparison, variation = "SD")
```

