---
title: "q2 example"
author: "Daniel Detore"
output: pdf_document
---

```{r}
VotingSurvey = data.frame(read.csv("VotingSurvey.csv"))
VotingSurvey$Vote = as.factor(VotingSurvey$Vote)
VotingSurvey$Gender = as.factor(VotingSurvey$Gender)
VotingSurvey$Party = as.factor(VotingSurvey$Party)
```

# 1
## 1.1
$H_0:$ 'Vote' and 'Gender' have no statistical association.

$H_a:$ 'Vote' and 'Gender' have some statistical association.

In this case the statistical association is whether voter's genders make them more
(or less) likely to vote.

## 1.2
```{r}
table = table(VotingSurvey$Vote, VotingSurvey$Gender); table
```

## 1.3

```{r}
barplot(table, legend=TRUE, beside=TRUE, main='Gender/Vote Barplot')
```
They are of different patterns. Females seem much more likely to vote and 
men are slightly less likely, but we should still test to see if these differences
are statistically relevant.

## 1.4
The value of each inner cell $e_{i,j} = \frac{o_{i,+}o_{+,j}}{o_{+,+}}$.
```{r}
o_ip <- rowSums(table)
o_pj <- colSums(table)
o_pp <- sum(table)
e <- outer(o_ip, o_pj)/o_pp
rownames(e) = c("No", "Yes")
colnames(e) = c("Female", "Male")
e
```
## 1.5
```{r}
chisq.test(table)
```
The testing statistic 
$$\chi^2 = \sum_{i=1}^r \sum_{j=1}^c \frac{(O_{i,j}-E_{i,j})^2}{E_{i,j}}$$ 
is observed as 
$$\sum_{i=1}^r \sum_{j=1}^c \frac{(o_{i,j}-e_{i,j})^2}{e_{i,j}}=1.76$$

The null distribution is $\chi^2_k$ where degrees of freedom 
$k=(r-1)(c-1) = (2-1)(2-1)=1$. 
The p-value is approximately 0.18.

If we take $\alpha = 0.1$ then we cannot reject $H_0$ because $\alpha < 0.18$, 
thus we must assume 'Vote' and 'Gender' have no statistical association.





## 2
## 2.1
$H_0:\ \mu_1 = \mu_2 = \mu_3$; i.e. the mean age of all parties are equal.

$H_a:$ The mean age of all parties are not all equal.

## 2.2
```{r}
boxplot(xlab = "Party", ylab = "Age", VotingSurvey$Age ~ VotingSurvey$Party, 
beside = TRUE, main = "Party/Age Boxplot")
```
The variances in age seem similar for all parties, but the mean for Rep looks slightly 
lower than the others.

## 2.3
```{r}
report = aov(Age ~ Party, VotingSurvey)
summary(report)
```

## 2.4

$$SSB = \sum_{i=1}^k n_i(\bar{X}_{i,.}-\bar{X}_{.,.})^2 = 127$$

$$SSE = \sum_{i=1}^k \sum_{j=1}^{n_i} (X_{i,j}-\bar{X}_{i,.})^2 = \sum_{i=1}^k(n_i-1)S_i^2 = 7429$$

$$SST = SSB+SSE = 7556$$

$$MSB = \frac{SSB}{k-1} = 63.4$$

$$MSE = \frac{SSE}{n-k} = 125.9$$

## 2.5
Our testing statistic $F=\frac{MSB}{MSE}=\frac{\frac{SSB}{k-1}}{\frac{SSE}{n-k}}$ 
is observed as $f=0.504$.
The null distribution is 
$\mathcal{F}_{k-1,n-k} = \mathcal{F}_{3-1,62-3}=\mathcal{F}_{2,59}$. 
The p-value is 0.607.

## 2.6
```{r}
summary(lm(Age ~ Party, data=VotingSurvey))$r.squared 
```
Since $\alpha = 0.1 < 0.607$, we cannot reject $H_0$,
thus we must assume the mean age of all parties are equal.

## 2.7
Using the box plot from 2.2, we may wat to  check if the mean age for Rep party is 
lower than the average mean age of Dem and Ind parties,
i.e. whether $\frac{1}{2}\mu_1 + \frac{1}{2}\mu_2 - \mu_3 = 0$.

## 2.8
```{r}
contrasts(VotingSurvey$Party) = c(1/2, 1/2, -1)
go = aov(Age ~ Party, VotingSurvey)
summary(go, split = list(Party = list(`mu3 vs 1/2 mu1 + 1/2 mu2` = 1)))
```
We have observed testing statistic $f = 0.997$ which gives p-value 0.322.
Since the p-value $0.322 > \alpha = 0.1$, we cannot reject $H_0$ and thus we must
that the mean age of the Rep party is equal to the average mean ages of the
Ind and Dem parties.

## 2.9
We can do multiple comparison using Fischer's Least Significant Difference.
```{r}
library(agricolae)
comparison = LSD.test(go, "Party", p.adj = "none")
comparison
```
And we can visualize the comparison as such:
```{r}
plot(comparison, variation = "SD")
```
Because these functions put all of the parties in the "a" group, we can conclude that
all parties' actual mean ages are equal.