---
title: "Midterm"
author: "Daniel Detore"
date: "March 13, 2025"
output: pdf_document
---

```{r}
library(readxl)
GpaGender <- read_excel("GpaGender.xls")
GPA <- GpaGender$GPA; GPA
n <- length(GPA); n
alpha <- 0.05
```

# Problem 1
```{r}
B_GPA <- ifelse(GPA > 7, 1, 0); B_GPA
phat <- sum(B_GPA)/length(B_GPA); phat
p0 <- 0.30
```
$H_0: p = 0.30$ with null distribution $N(0,1)$

$H_a: p > 0.30$

We are testing one population proportion, so we use observation $z$ of $Z$ given by:
```{r}
z <- sqrt(n)*(phat - p0) / sqrt(p0 * (1 - p0)); z
```
which makes the p-value for the left-tail test:
```{r}
pnorm(z)
```
which is greater than our $\alpha = 0.05$. We cannot disprove $H_0$.

Using the R function:
```{r}
prop.test(sum(B_GPA), length(B_GPA), p = p0, alternative = c("less"), correct = FALSE)
```
Where $p = 1 > \alpha = 0.05$. We cannot disprove $H_0$.

# Problem 2
```{r}
M_GPA <- GpaGender[which(GpaGender$Gender == 2), "GPA"]; M_GPA
F_GPA <- GpaGender[which(GpaGender$Gender == 1), "GPA"]; F_GPA

# Male summary:
summary(M_GPA)
# Female summary:
summary(F_GPA)
```

# Problem 3
```{r}
both_GPA <- c(M_GPA, F_GPA); both_GPA
boxplot(both_GPA, names = c("male", "female"), horizontal = TRUE)
```

# Problem 4
```{r}
t.test(M_GPA, conf.level = 0.9)
t.test(F_GPA, conf.level = 0.9)
```
90% Confidence interval for males is $[6.713815, 7.849462]$; for females it is $[7.171981, 8.221116]$.

# Problem 5
```{r}
M_GPA <- GpaGender[which(GpaGender$Gender == 2), "GPA"]$GPA; M_GPA
F_GPA <- GpaGender[which(GpaGender$Gender == 1), "GPA"]$GPA; F_GPA
```
$H_0: \sigma^2_M = \sigma^2_F$ where $H_0 \sim F_{n_M-1,\ n_F-1} \equiv F_{46,\ 30}$

Let's check our sample variances:

```{r}
SSM <- var(M_GPA); SSM
SSF <- var(F_GPA); SSF
```

Since $S_M^2 > S_F^2$, $H_a: \sigma^2_M > \sigma^2_F$

Our observed $F = f =$
```{r}
f <- SSM/SSF; f
```
and gives us p-value
```{r}
1 - pf(f, length(M_GPA) - 1, length(F_GPA) - 1)
```
which is less than $\alpha = 0.05$. We can reject $H_0$ and assume $\sigma^2_M > \sigma^2_F$.
Using the R function:
```{r}
var.test(M_GPA, F_GPA, ratio = 1, alternative = c("greater"), conf.level = 1 - alpha)
```

# Problem 6
We know $\sigma^2_M \neq \sigma^2_F$ and we know neither value. So we will do a general two-sample T-test.

$H_0: \mu_M = \mu_F$

$H_a: \mu_M \neq \mu_F$

We calculate our observed $T = t$ where
```{r}
xbar <- mean(M_GPA); xbar
ybar <- mean(F_GPA); ybar
n1 <- length(M_GPA); n1
n2 <- length(F_GPA); n2
t <- (xbar - ybar)/sqrt(SSM/n1 + SSF/n2); t
```
and get a p-value with
```{r}
k <- (SSM/n1+SSF/n2)^2/((SSM/n1)^2/(n1-1)+(SSF/n2)^2/(n2-1))
k <- ceiling(k); k
2 * pt(-abs(t), k)
```
which is greater than 0.05, so we cannot reject $H_0$, so $\mu_M = \mu_F$.

Using the R function:
```{r}
t.test(M_GPA, F_GPA, alternative = c("two.sided"), paired = FALSE, var.equal = FALSE)
```